What covert verbal actions tell us about the interplay between language, action, and perception

### 2021-06-08 manuscript.tex
minor updates

In this perspective, we examine some theoretical and experimental consequences of considering covert verbal actions as inhibited verbal actions. We explore two main themes. First, we discuss the plausible origins of the multisensory (e.g., auditory, kinaesthetic) content of covert verbal actions. Second, we discuss the under-explored role of inhibitory mechanisms during motor imagery and covert verbal actions more specifically. By bridging together recent results from motor imagery, motor inhibition, and inner speech research, we highlight some novel and possibly fruitful lines of research.

@@

Where does the sensory content of covert verbal actions come from? How does motor learning influence the multisensory content of covert verbal actions? In other words, how do the motor and perceptual system interact during covert verbal actions? In the first subsection, we discuss two mechanisms that may be responsible for providing the sensory content of covert verbal actions and their respective involvement in different forms of covert speech. In the second subsection, we discuss an exciting avenue for investigating the influence of motor learning on the sensory content of covert verbal actions, namely, the imagery of totally novel (i.e., never executed) verbal actions.

@@

The dual stream prediction model \citep{tian_mental_2012, tian_effect_2013, tian_mental_2016}... Direct simulation (attention-guided memory retrieval process) or indirect simulation (motor simulation process) \citep{tian_mental_2012, tian_effect_2013, li_corollary_2020, ma_distinct_2019}? The dual stream prediction model \citep{tian_mental_2016}... \citep[similar to the distinction between the prediction-by-simulation or prediction-by-association mechanisms of][]{pickering_integrated_2013}...

The balance between these two mechanisms may depend by the precise instructions given to the participants. For instance, depending on whether participants are instructed to "imagine speaking" or to "imagine hearing" speech (similar to the distinction between the "inner ear" and the "inner voice", e.g., \cite{smith_subvocalization_1992})... Congruent with this hypothesis, it has been shown that inner speaking and inner hearing have distinct MEG correlates and effects on a subsequent /ba/-/ta/ categorisation task \citep{ma_distinct_2019}...

In the absence of non-ambiguous instructions, and in line with \cite{tian_mental_2012}, we suggest that the balance between these two mechanisms may also depend on situational (e.g., surrounding noise) and intrinsic (e.g., expertise) characteristics. For instance... In addition to \cite{tian_mental_2012}, we suggest that a common currency determining the use of one of these mechanisms is the computational cost/resource of each alternative. To clarify, we borrow the concept of memoisation as applied to cognition and mental imagery by \cite{dasgupta_memory_2021}. In this context, memory can be considered as a mechanism to facilitate computational reuse. In the same way... efficient cognition may rely on stored past computation results rather than reinventing the wheel every time. In the context of motor and speech imagery... In other words, situational (extrinsic) and individual (intrinsic) characteristics jointly determine the computational cost (or equivalently, the available computational resources), which in turn determines the balance between the simulation and association mechanisms. Novel and/or difficult tasks may rely more on the simulation mechanism whereas well known tasks may rely more on associative mechanisms... This idea is supported by several studies, showing an increase in facial EMG activity during the reading of difficult text \cite{babar} or while performing difficult mental arithmetic tasks \cite{babar}. See also the discussion in \citep{nalborczyk_understanding_2019-1, nalborczyk_re-analysing_2020}.

Memoisation: avoid redundant computation by storing computational results and reusing them later... functions call are intercepted with by a memoiser the inspects the previous calls of a function and its outputs. If a function has already been called with the same input, then the previously computed output is retrieved and reused... memoisation can be made more flexible (generalisable) by using \textit{interpolation} between cached outputs... for instance, in the context of motor control, reaching a goal (e.g., a cup of tea) can be decomposed in motor primitives, where input-output mapping can be retrieved from a memo table and/or interpolated from sufficiently similar stored mappings... or as put by \cite{dasgupta_memory_2021}, the "shift from algorithm to memory"... this shift is consistent with studies showing a lesser involvement of the motor system for well known imagined actions... Depending on whether the task needs to form "precise" sensory representations and whether we have the time resources to do it... Training studies also suggest that we rely more and more on domain-general (e.g., memory-based) processes with expertise \citep[e.g.,][]{tarr_mental_1989, jolicoeur_time_1985}.

@@

How can we imagine totally novel (i.e., never executed) actions? Broadly speaking, the simulationnist framework suggests that the multisensory experience of motor imagery would result from the simulation of the corresponding motor action, reusing internal models developed for the control of overt actions. In other words, internal models developed for the control of actions would then be used to simulate the sensory consequences of covert actions. Notice that these internal models are generative models, which means that they can a priori be used to compute (predict) the sensory consequences of any action that can be parsed as a (possibly supramodal) goal to an inverse internal model. In other words, this framework makes it possible for novel actions to be simulated internally.\footnote{\cite{mulder_role_2004} assessed whether the mental practice of a totally novel movement (abduction of the big toe of the dominant right foot) resulted in better overt execution over the course of several weeks of practice. They observed that although mental practise did improve execution in the group of participants that already had some experience with the task at baseline, mental practise did not lead to statistically significant improvements in execution in the group of participants that had zero experience with the task at baseline. However, a Bayesian reanalysis of their results suggests that the evidence in favour of the null hypothesis was rather inconclusive ($\text{BF}_{01} = 1.92$, using a weakly informative Cauchy prior on the standardised effect size for the alternative hypothesis, with $r = 1$).}

A prediction of this putative mechanism is that the imagination of novel actions, and more precisely, the predicted sensory consequences of novel actions, may be "biased" or "constrained" by the repertoire of already-known actions. This prediction could be assessed by asking participants to imagine a novel action and subsequently checking whether this improves the execution of an intermediate (interpolated) novel action in the articulatory space. The same logic could be applied to the covert production of speech with someone else’s voice. How can we imagine speaking with the voice of our relatives? If it is based on simulators (i.e., on internal models), then the covert production of speech with someone else’s voice should possess sensory properties that are somehow "biased" or "constrained" by the properties of our own sensory and motor systems. Taking as an illustration the auditory sensations of speech, the subjective auditory sensation of covert speech with someone else’s voice should be situated in a somehow intermediate location in the auditory (e.g., formantic) space between the acoustic properties of our own voice and the acoustic properties of someone else voice.

This prediction could be assessed experimentally by... the prediction-by-simulation and the prediction-by-association accounts could be distinguished by constraining the motor system and assessing whether this influences the internal sensory representation of the produced speech. If it does...

@@

The proposal that overt and covert actions share common neural circuits is faced with a serious problem. If the neural circuits used for the control of overt actions are reused for covert actions, how can covert actions not lead to execution? This puzzle, coined as \textit{the problem of inhibition} by \cite{jeannerod_neural_2001}, can be rephrased as follows: given the putative role of the motor system in providing the multisensory content of motor imagery, how is it possible for motor imagery not to lead to motor execution? In the following, we briefly summarise the evidence in favour of the view of covert verbal actions as inhibited verbal actions. Then, we relate this view with the presumed progressive internalisation of speech during childhood. Finally, we outline some working hypotheses regarding the cognitive and neural implementation of these inhibitory mechanisms.

@@

First and foremost, we need to make a distinction between at least two different types of inhibition: the inhibition of physical response (e.g., planning to raise the arm and finally not doing it), and the idea of cognitive inhibition, defined as "the stopping or overriding of a mental process, in whole or in part, with or without intention" \citep{gorfein_concept_2007}. Here we are concerned with the former, which can be defined broadly as the withholding, suppression, or overriding of an inappropriate, prepotent, or unwanted motor response \citep{aron_neural_2007, oshea_go_2018}. To be even more precise, \cite{ridderinkhof_dont_2014} described the concept of response inhibition on three continuous dimensions: intentionality, premeditation, specificity. Inhibition can be employed with more or less intentionality (intentional vs. reactive inhibition), it can planned ahead or employed in the moment (early vs. late inhibition), and it can be more or less specifically (global inhibition versus effector-specific or action-specific inhibition).

With regards to covert verbal actions specifically, we hypothesise that they involve an intentional (we know we want to produce motor imagery rather than overt execution) but automatic (we do not explicitly think about not producing movements), planned ahead (proactive), and possibly global form of response inhibition... or maybe both global and effector-specific, cf. the results of \cite{rieger_inhibition_2017}... Notice that the type of motor inhibition that may be at play during motor imagery is still different from the "proactive inhibition" in the motor inhibition literature, in the sense that while doing motor imagery, the participants is not explicitly asked to not produce an action. Instead, they are asked to produce the action \textit{covertly}, which (only implicitly) implies that it should be executed overtly.

@@

% Summary of the main proposition (i.e., considering covert verbal actions as inhibited verbal actions) and its consequences on the relation between perception and action, theoretical and experimental perspectives...

To sum up, we explored the analogy between covert verbal actions such as covert speech and other forms of motor imagery. This led us to a discussion of two main themes, that is, the question of how the sensory content of motor imagery is produced, and the question of how is it possible for motor imagery not to lead to motor execution. To sketch an answer to these questions, we connected results from the motor imagery, motor inhibition, and inner speech research traditions.

We suggested that (section 1 propositions)... and (section 2 propositions)... Taken together, these propositions pave the way for several lines of research that should...

>>>>

In this perspective, we examine some theoretical and experimental consequences of considering covert verbal actions as other forms of imagined actions. We explore two main themes. First, we discuss the plausible origins of the multisensory (e.g., auditory, kinaesthetic) content of covert verbal actions. Second, we discuss the under-explored role of inhibitory mechanisms during motor imagery and covert verbal actions more specifically. By bridging together recent results from the motor imagery, motor inhibition, and covert speech literature, we highlight some novel and possibly fruitful lines of research.

@@

Where does the sensory content of covert verbal actions come from? How does motor learning shape these sensory percepts? In other words, how do the motor and perceptual system interact during covert verbal actions? In the first subsection, we discuss two mechanisms that may be responsible for providing the sensory content of covert verbal actions and their respective involvement in different forms of covert speech. In the second subsection, we discuss an exciting avenue for investigating the influence of motor learning on the sensory content of covert verbal actions, namely, the imagery of totally novel (i.e., never executed) verbal actions.

@@

The dual stream prediction model \citep{tian_mental_2012, tian_effect_2013, tian_mental_2016}... Direct simulation (attention-guided memory retrieval process) or indirect simulation (motor simulation process) \citep{tian_mental_2012, tian_effect_2013, li_corollary_2020, ma_distinct_2019}? The dual stream prediction model \citep{tian_mental_2016}... \citep[this proposal shares similarities with the distinction between the prediction-by-simulation or prediction-by-association mechanisms of][]{pickering_integrated_2013}.

The balance between these two mechanisms may depend on the precise instructions given to the participants, which may clue them to produce different form of covert speech. For instance, depending on whether participants are instructed to "imagine speaking" or to "imagine hearing" speech \citep[see also the distinction between the "inner ear" and the "inner voice", e.g.,][]{smith_subvocalization_1992}... Congruent with this hypothesis, it has been shown that inner speaking and inner hearing have distinct MEG correlates and effects on a subsequent /ba/-/ta/ categorisation task \citep{ma_distinct_2019}. Moreover...

In the absence of non-ambiguous instructions, and in line with \cite{tian_mental_2012}, we suggest that the balance between these two mechanisms may also depend on situational (e.g., surrounding noise) and intrinsic (e.g., expertise) characteristics. For instance... In addition to \cite{tian_mental_2012}, we suggest that a common currency determining the use of one of these mechanisms is the computational cost of (or equivalently, the computational resources available for) each alternative. To clarify, we borrow the concept of memoisation as applied to cognition and mental imagery by \cite{dasgupta_memory_2021}.\footnote{Memoisation is a programming technique used to speed-up algorithms or programs. It avoids redundant computation by storing computational results and reusing them later. When calling a function (where a function can be a motor primitive such as elevating the arm), the function call is intercepted by a memoiser that inspects the previous calls of a function and its outputs. If a function has already been called with the same input, then the previously computed output is retrieved and reused.} In this context, memory can be considered as a mechanism to facilitate computational reuse. In the same way efficient programming may rely on computational reuse, efficient cognition may rely on stored past computation results rather than reinventing the wheel every time. In the context of motor and speech imagery, the memory-retrieval predictions stream would be the memoised version of the simulation-estimation prediction stream, where motor-to-sensory mappings are (party or completely) retrieved from memory, instead of being computed again...

In other words, situational (extrinsic) and individual (intrinsic) characteristics jointly determine the computational cost (or equivalently, the available computational resources), which in turn determines the balance between the simulation and association mechanisms. For instance, we hypothesise that novel and/or difficult tasks (which are both computationally more expensive, ceteris paribus) may rely more on the simulation mechanism, whereas well known and/or easy tasks may rely more on associative mechanisms. This idea is supported by several studies showing a greater increase in facial EMG activity during the reading of difficult text or while performing difficult mental arithmetic tasks as compared to easier tasks \citep[e.g.,][]{faaborg-andersen_electromyography_1958, sokolov_inner_1972}, suggesting a greater involvement of the speech motor system \citep[and/or a lesser involvement of inhibitory mechanisms, see also the discussion in][]{nalborczyk_understanding_2019-1, nalborczyk_re-analysing_2020}. This is consistent with results from training studies suggesting that we rely more and more on domain-general (e.g., memory-based) processes with expertise \citep[e.g.,][]{tarr_mental_1989, jolicoeur_time_1985}.

% Memoisation: avoid redundant computation by storing computational results and reusing them later... functions call are intercepted with by a memoiser the inspects the previous calls of a function and its outputs. If a function has already been called with the same input, then the previously computed output is retrieved and reused... memoisation can be made more flexible (generalisable) by using \textit{interpolation} between cached outputs... for instance, in the context of motor control, reaching a goal (e.g., a cup of tea) can be decomposed in motor primitives, where input-output mapping can be retrieved from a memo table and/or interpolated from sufficiently similar stored mappings... or as put by \cite{dasgupta_memory_2021}, the "shift from algorithm to memory"... this shift is consistent with studies showing a lesser involvement of the motor system for well known imagined actions... Depending on whether the task needs to form "precise" sensory representations and whether we have the time resources to do it... Training studies also suggest that we rely more and more on domain-general (e.g., memory-based) processes with expertise \citep[e.g.,][]{tarr_mental_1989, jolicoeur_time_1985}.

@@

If, as suggested in the previous section, the sensory content of covert verbal actions is (at least in some situations) provided by a motor simulation process, how can we imagine totally novel (i.e., never executed) actions? Broadly speaking, the simulationnist framework suggests that the multisensory experience of motor imagery results from the simulation of the corresponding motor action, reusing internal models developed for the control of overt actions. In other words, internal models developed for the control of actions would then be used to simulate the sensory consequences of covert actions. Crucially, these internal models are generative models, which means that they can be used to compute (i.e., predict) the sensory consequences of any action that can be parsed as a (possibly supramodal) goal to an inverse internal model. In other words, this framework makes it possible for novel actions to be simulated internally.\footnote{\cite{mulder_role_2004} assessed whether the mental practice of a totally novel movement (abduction of the big toe of the dominant right foot) resulted in better overt execution over the course of several weeks of practice. They observed that although mental practise did improve execution in the group of participants that already had some experience with the task at baseline, mental practise did not lead to statistically significant improvements in execution in the group of participants that had zero experience with the task at baseline. However, a Bayesian reanalysis of their results suggests that the evidence in favour of the null hypothesis was rather inconclusive ($\text{BF}_{01} = 1.92$, using a weakly informative Cauchy prior on the standardised effect size for the alternative hypothesis, with $r = 1$).}

A prediction of this putative mechanism is that the imagination of novel actions, and more precisely, the predicted sensory consequences of novel actions, may be "biased" or "constrained" by the repertoire of already-known actions. For instance, when asked to imagine playing table tennis, a tennis player may overestimate movements' amplitude and duration. Translated to speech production, this prediction could be assessed by asking participants to imagine a novel action and subsequently checking whether this improves the execution of an intermediate (interpolated) novel action in the articulatory space. The same logic could be applied to the covert production of speech with someone else's voice. How can we imagine speaking with the voice of our relatives? If it is based on simulators (i.e., on internal models), then the covert production of speech with someone else's voice should possess sensory properties that are somehow "biased" or "constrained" by the properties of our own sensory and motor systems. Taking as an illustration the auditory sensations of speech, the subjective auditory sensation of covert speech with someone else's voice should be situated in a somehow intermediate location in the auditory (e.g., formantic) space, between the acoustic properties of our own voice and the acoustic properties of someone else voice.

% This prediction could be assessed experimentally by... the prediction-by-simulation and the prediction-by-association accounts could be distinguished by constraining the motor system and assessing whether this influences the internal sensory representation of the produced speech. If it does...

This prediction could be assessed experimentally by using the sensory attenuation effect observed during covert speech production. Auditory stimuli elicit an electrophysiological brain response (the auditory-evoked potential) with a characteristic N1 component. This N1 component has been shown to be reduced when the presentation of an auditory stimulus is synchronised with the self-production of a phonetically-matched vocalisation (an effect called N1-suppression). More interestingly, \cite{whitford_neurophysiological_2017} shown that the covert production of a phoneme also resulted in sensory attenuation (as assessed by N1-suppression) if the content of the imagined phoneme matched the content of the audible phoneme. Therefore, we could describe the subjective auditory sensation of covert speaking with someone else’s voice by manipulating the acoustic (e.g., formantic) properties of the auditory stimulus that is presented to the participant to find the stimulus that produces the greatest sensory attenuation (as assessed by N1-suppression). In other words, sensory attenuation should be the strongest when it matches the auditory representation of the speaker. If the auditory percepts generated when covert speaking with the voice of someone else is "constrained" by our own motor system, then this point of greatest sensory attenuation should occur in an intermediate location in the feature (e.g., formantic) space.

@@

The proposal that overt and covert actions share common neural circuits is faced with a serious problem. If the neural circuits used for the control of overt actions are reused for covert actions, how can covert actions not lead to execution? This puzzle, coined as \textit{the problem of inhibition} by \cite{jeannerod_neural_2001}, can be rephrased as follows: given the putative role of the motor system in providing the multisensory content of motor imagery, how is it possible for motor imagery not to lead to motor execution? In the following, we briefly summarise the evidence in favour of the view of covert verbal actions as inhibited verbal actions. Then, we outline some working hypotheses regarding the cognitive and neural implementation of these inhibitory mechanisms. Finally, we relate this view with the presumed progressive internalisation of speech during childhood.

@@

First and foremost, we need to make a distinction between at least two different types of inhibition: the inhibition of physical response (e.g., planning to raise the arm and finally not doing it), and the idea of cognitive inhibition, defined as "the stopping or overriding of a mental process, in whole or in part, with or without intention" \citep{gorfein_concept_2007}. Here, we are concerned with the former, which can be defined broadly as the withholding, suppression, or overriding of an inappropriate, prepotent, or unwanted motor response \citep{aron_neural_2007, oshea_go_2018}. To be even more precise, \cite{ridderinkhof_dont_2014} described the concept of response inhibition on three continuous dimensions: intentionality, premeditation, specificity. Inhibition can be employed with more or less intentionality (intentional vs. reactive inhibition), it can planned ahead or employed in the moment (early vs. late inhibition), and it can be more or less specifically (global inhibition versus effector-specific or action-specific inhibition).

With regards to covert verbal actions specifically, we hypothesise that they involve an intentional (we know we want to produce motor imagery rather than overt execution) but automatic (we do not explicitly think about not producing movements), planned ahead (proactive), and possibly global form of response inhibition. Or maybe both global and effector-specific, cf. the results of \cite{rieger_inhibition_2017}... Notice that the type of motor inhibition that may be at play during motor imagery is still different from the "proactive inhibition" in the motor inhibition literature, in the sense that while doing motor imagery, the participants is not explicitly asked to not produce an action. Instead, they are asked to produce the action \textit{covertly}, which (only implicitly) implies that it should be executed overtly.

@@

% Summary of the main proposition (i.e., considering covert verbal actions as inhibited verbal actions) and its consequences on the relation between perception and action + theoretical and experimental perspectives.

To sum up, we explored (some of) the theoretical and experimental perspectives that emerge when considering covert verbal actions as other forms of imagined actions. This led us to a discussion of two main themes. First, we discussed the plausible origins two of the sensory content of convert verbal actions is produced. Second, we asked how is it possible for motor imagery not to lead to motor execution. To sketch an answer to these questions, we connected results from the motor imagery, motor inhibition, and inner speech research traditions.

Regarding the origin of the sensory experience of covert verbal actions, we discussed the dual stream prediction model \citep{tian_mental_2012, tian_effect_2013, tian_mental_2016}, which suggests that these sensory percepts may be provided either by a motor-simulation-based process or by a memory-based process. We suggested that the balance between these two mechanisms, in addition to the type of covert speech that is prompted by the tasks instructions, may also be determined by the available computational resources (or equivalently, the cost of the task, with novel or more difficult tasks being more costly). More precisely...

Regarding the role of general-purpose inhibitory mechanisms during the production of covert verbal actions, we suggested that these may be similar to the inhibitory network responsible for proactive response inhibition and summarised some propositions from this literature. We related the development of response inhibition abilities in childhood development with the purported internalisation of private speech around the same period. From the response inhibition perspective, the internalisation of speech from overt to covert speech may essentially be considered as "learning not to produce speech" (i.e., learning pairs of internal models that will be used to simulate speech motor-to-sensory mappings).

Taken together, these propositions pave the way for several lines of research that should... conclusion...
